---
title: "DataMining_Project_2022_23"
author: "Joel Pablos Martin"
date: "2023-04-28"
output: html_document
---

```{r}
knitr::opts_knit$set(root.dir ="/srv/data/joelpm/Code")
```

```{r, eval=FALSE}
# To convert the .rmd to .R
knitr::purl("/srv/data/joelpm/Code/DataMining_Project_2022_23.rmd", documentation=2)
```


```{r}
# install.packages("vctrs")
# install.packages('Seurat')
# BiocManager::install("DESeq2")
# install.packages('Matrix')
# BiocManager::install(c('Rsamtools', 'ChIPseeker', 'monocle','RcisTarget', 'RCy3', 'clusterProfiler'))
# install.packages("devtools")
# devtools::install_github("jiang-junyao/IReNA")
# install.packages("scTenifoldNet")
# install.packages("remotes")
# remotes::install_github("aertslab/SCENIC")
# install.packages("igraph")
# BiocManager::install("clusterProfiler")
library(vctrs)
library(Seurat)
library(DESeq2)
library(Matrix)
library(devtools)
library(IReNA)
library(scTenifoldNet)
library(SCENIC)
library(igraph)
library(clusterProfiler)
```


```{r}
knitr::opts_knit$set(root.dir ="/srv/data/joelpm/Code")
# https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE192740
```

# Load dataset
```{r}
#/home/joelpm/Projects/Data/ATLAS/MouseStSt/All_cells/rawData_mouseStSt/countTable_mouseStSt
All_expression_10X <- Read10X("/srv/data/joelpm/Data/ATLAS/MouseStSt/All_cells/countTable_mouseStSt",
                              gene.column = 1,
                              cell.column = 1,
                              unique.features = TRUE,
                              strip.suffix = TRUE)

All_expression_10X[1:10,1:10]
dim(All_expression_10X)
```



We load the annotation metadata
```{r}
# Load your annotations CSV file (previously adapted in scanpy)
annot_all <- read.csv("/srv/data/joelpm/Data/ATLAS/MouseStSt/All_cells/annot_mouseStStAll_withStellateandLSECsandMono.csv", header = TRUE, stringsAsFactors = FALSE)
annot_all <- annot_all[order(annot_all$annot),]
annot_all

# Remove cells that did not pass the filtering from the Liver cell Atlas publication
All_cells <- colnames(All_expression_10X)
All_cells_filt <- All_cells[All_cells %in% annot_all$cell]
All_expression_filt <- All_expression_10X[,All_cells %in% annot_all$cell]
dim(All_expression_10X)
dim(All_expression_filt)
length(annot_all$cell)

# reorder the rows to match the order
annot_all <- annot_all[match(All_cells_filt, annot_all$cell), ]
rownames(annot_all) <- seq(1,nrow(annot_all))
annot_all
write.table(annot_all, "/srv/data/joelpm/Data/ATLAS/MouseStSt/All_cells/annot_mouseStStAll_withStellateandLSECsandMono_sortSeurat.csv", sep=",", row.names=FALSE, col.names = TRUE, quote=FALSE)
```

```{r}
# Create Seurat Object
All_Seurat <- CreateSeuratObject(counts = All_expression_filt, project = "AllLiverCells", min.cells = 3, min.features = 200)

# Normalization of the data 
All_Seurat <- NormalizeData(All_Seurat, normalization.method = "LogNormalize", scale.factor = 10000)

# Load annotation in Seurat object
All_Seurat@meta.data <- annot_all
rownames(All_Seurat@meta.data) <- All_Seurat@meta.data$cell
Idents(All_Seurat) = All_Seurat@meta.data[["annot"]]
All_Seurat <- AddMetaData(All_Seurat, annot_all)
```


```{r}
levels(All_Seurat)
```

```{r, echo=FALSE, eval=FALSE}
# We create a S4 object to store all the expression data by populations
# Extract unique cell types
cell_types <- unique(annot_all$annot)

# Initialize an empty character vector to store slot definitions
slot_defs <- character(length = length(cell_types))
slot_set <- character(length = length(cell_types))

# Loop over cell types and create a slot definition for each one
cell_names <- c()
for (i in seq_along(cell_types)) {
  cellname <- gsub(" ", "_", cell_types[i])
  cellname <- gsub("&", "and", cellname)
  cellname <- gsub("-", "_", cellname)
  cellname <- gsub("\\.", "", cellname)
  cell_names <- c(cell_names,cellname)
  slot_defs[i] <- paste0(cellname, " = 'list',")
  slot_set[i] <- paste0(cellname, " = list(),")
}

# Combine slot definitions into a single string
slot_defs <- paste0(slot_defs, collapse = " ")
slot_set <- paste0(slot_set, collapse = " ")
slot_defs <- sub(",\\s*$", "", slot_defs)
slot_set <- sub(",\\s*$", "", slot_set)

# Create the S4 class with the dynamic slot definitions
class_def <- paste0("setClass('MyClass_NICHE_all', slots = c(", slot_defs, "))")
class_set <- paste0("NICHE_specific <- new('MyClass_NICHE_all',  ", slot_set, ")")
eval(parse(text = class_def))
eval(parse(text = class_set))

pseudobulk_all <- data.frame(matrix(NA, nrow = nrow(All_Seurat), ncol = 0))
for (c in seq(1,length(cell_types))){
  pob.cell <- annot_all$cell[annot_all$annot == cell_types[c]]
  pob.mx <- All_Seurat[,colnames(All_Seurat) %in% pob.cell]
  pob.mxnorm <- All_Seurat_filtnorm[,colnames(All_Seurat_filtnorm) %in% pob.cell]
  pob.bulk <- rowSums(pob.mx)
  df.bulk <- data.frame(pob.bulk)
  colnames(df.bulk) <- cell_types[c]
  pseudobulk_all <- cbind(pseudobulk_all, df.bulk)
  slot(NICHE_specific, cell_names[c])[["Seurat_matrix"]] <- pob.mx
  slot(NICHE_specific, cell_names[c])[["Seurat_matrix_normalized"]] <- pob.mxnorm
  slot(NICHE_specific, cell_names[c])[["pseudobulk_raw"]] <- pob.bulk
}
pseudobulk_all

rm(pob.cell, pob.mx, pob.mxnorm, c, i, cell_names, cellname, slot_defs, slot_set, class_def, class_set)
```


# scTenifoldNet Monocyte vs Kupffer Cell
https://pubmed.ncbi.nlm.nih.gov/31587991/

## 1. Load the datasets
```{r}
set.seed(1127)
# Monocytes
SeuratNorm.Mono <- GetAssayData(All_Seurat,slot = 'data')[,which(Idents(All_Seurat) == "Monocytes")]
dim(SeuratNorm.Mono)
## filtering to just keep genes with expression non-zero in at least 5% of the cells
prop_expr.Mono <- apply(SeuratNorm.Mono > 0, 1, mean)
keep_genes.Mono <- names(prop_expr.Mono[prop_expr.Mono >= 0.05])

# Kupffer Cells
SeuratNorm.KC <- GetAssayData(All_Seurat,slot = 'data')[,which(Idents(All_Seurat) == "Kupffer cells")]
dim(SeuratNorm.KC)
## filtering to just keep genes with expression non-zero in at least 5% of the cells
prop_expr.KC <- apply(SeuratNorm.KC > 0, 1, mean)
keep_genes.KC <- names(prop_expr.KC[prop_expr.KC >= 0.05])

# Union of genes expressed in both cell types and filtering
union.genes <- unique(c(keep_genes.Mono, keep_genes.KC))
mtGenes <- rownames(SeuratNorm.Mono)[grepl('^mt-', rownames(SeuratNorm.Mono), ignore.case = TRUE)]
keyGenes <- c("Clec4f","Spic","Nr1h3","Spi1","Rxra", "Rbpj")
mtGenes <- NULL
keyGenes <- NULL
# union.genes <- sample(unique(keep_genes.Mono,keep_genes.KC), size = 100, replace=F) # genes
SeuratNorm.Mono <- SeuratNorm.Mono[row.names(SeuratNorm.Mono) %in% c(union.genes, mtGenes, keyGenes),]
SeuratNorm.KC <- SeuratNorm.KC[row.names(SeuratNorm.KC) %in% c(union.genes, mtGenes, keyGenes),]

#subsampling to test faster
#SeuratNorm.Mono <- SeuratNorm.Mono[, sample(colnames(SeuratNorm.Mono), size = 2000, replace=F)] # cells

#subsampling to test faster
SeuratNorm.KC <- SeuratNorm.KC[, sample(colnames(SeuratNorm.KC), size = ncol(SeuratNorm.Mono), replace=F)] # cells

dim(SeuratNorm.Mono)
dim(SeuratNorm.KC)

keyGenes <- c("Clec4f","Spic","Nr1h3","Spi1","Rxra", "Rbpj")
keyGenes %in% rownames(SeuratNorm.Mono)
keyGenes %in% rownames(SeuratNorm.KC)
```

## 2. scTenifoldNet

```{r}
mx.SeuratNorm.Mono <- as.matrix(SeuratNorm.Mono)
mx.SeuratNorm.KC <- as.matrix(SeuratNorm.KC)
scTFNet.MonovsKC <- scTenifoldNet(X = mx.SeuratNorm.Mono, Y = mx.SeuratNorm.KC,
                          nc_nNet = 10, nc_nCells = 500,
                          td_K = 3, qc_minLibSize = 30,
                          qc = FALSE, nCores = 20)
scTFNet.MonovsKC
setwd("/srv/data/joelpm/Code")
# save(scTFNet.MonovsKC, ascii=FALSE, file="../Data/DataMining/scTFNet.MonovsKC.bin")
load("../Data/DataMining/scTFNet.MonovsKC_20230512.bin")
```

## 3. Weight-averaged denoised gene regulatory networks
```{r}
# Network of Monocytes
igraph.Mono <- igraph::graph_from_adjacency_matrix(adjmatrix = scTFNet.MonovsKC$tensorNetworks$X, weighted = TRUE)
igraph.Mono
plot(igraph.Mono)
# Network of Kupffer Cells
igraph.KC <- igraph::graph_from_adjacency_matrix(adjmatrix = scTFNet.MonovsKC$tensorNetworks$Y, weighted = TRUE)
igraph.KC
plot(igraph.KC)
```

## 4. Manifold alignment
In the manifold alignment, there are two rows for each gene, one for each sample:
```{r}
scTFNet.MonovsKC$manifoldAlignment[c('X_Nr1h3', 'y_Nr1h3'),]
```
Each pair of coordinates (representing a gene in each sample) is given a Euclidean distance, and the sum of the distances is used to conduct the differential regulation test.
```{r}
dist(scTFNet.MonovsKC$manifoldAlignment[c('X_Nr1h3', 'y_Nr1h3'),])
dist(scTFNet.MonovsKC$manifoldAlignment[c('X_Clec4f', 'y_Clec4f'),])
```

## 5. Manifold alignment distances as the basis for differential regulation
```{r}
diffReg.MonovsKC <- scTFNet.MonovsKC$diffRegulation
diffReg.MonovsKC <- diffReg.MonovsKC[order(diffReg.MonovsKC$p.adj, -diffReg.MonovsKC$distance, -diffReg.MonovsKC$FC),]
rownames(diffReg.MonovsKC) <- seq(1,nrow(diffReg.MonovsKC))
head(diffReg.MonovsKC, n = 100)
diffReg.MonovsKC[diffReg.MonovsKC$gene == "Nr1h3",]
diffReg.MonovsKC[diffReg.MonovsKC$gene == "Rbpj",]

write.table(diffReg.MonovsKC, "../Data/DataMining/scTFNet_MonovsKC_diffRegulation.tsv", sep="\t", row.names=FALSE, col.names = TRUE, quote=FALSE)
```
## Plotting the results
```{r}
library(ggplot2)
library(ggrepel)

diffReg.MonovsKC

top_labels <- diffReg.MonovsKC[diffReg.MonovsKC$p.adj < 0.05,]

plot_scTFNet <- ggplot(data = diffReg.MonovsKC, aes(x = FC, y = distance)) +
  geom_point(aes(size = p.adj, fill = ifelse(p.adj > 0.05, "deepskyblue3",
                                                    ifelse(p.adj <= 0.05, "red"))),
             shape = 21, color = "black" , alpha = 0.7, stroke = 0.2) +
  geom_text_repel(data = subset(top_labels), aes(label = gene), color = "black",
                  box.padding = 0.5, max.overlaps = Inf, force = 0.5,
                  segment.linetype = 1, segment.size = 0.2) +
  labs(x = "Log2 fold change in gene expression",
       y = "Distance",
       title = paste0("Monocyte to Kupffer cell: scTenifoldNet")) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_size_continuous(name = "P-value adjusted", range = c(4, 0)) +
  scale_fill_manual(values = c("deepskyblue3" = "deepskyblue3", "red" = "red")) +
  coord_cartesian(ylim = c(min(diffReg.MonovsKC$distance), 1))

plot_scTFNet
```

```{r}
library(ggplot2)
library(ggrepel)

top_labels <- diffReg.MonovsKC[diffReg.MonovsKC$p.adj < 0.05,]

plot_scTFNet <- ggplot(data = diffReg.MonovsKC, aes(x = FC, y = distance)) +
  geom_point(aes(size = p.adj, fill = ifelse(p.adj > 0.05, "deepskyblue3", "red")),
             shape = 21, color = "black", alpha = 0.7, stroke = 0.2) +
  geom_text_repel(data = subset(top_labels), aes(label = gene), color = "black",
                  box.padding = 0.5, max.overlaps = Inf, force = 0.5,
                  segment.linetype = 1, segment.size = 0.2) +
  labs(x = "Log2 fold change in gene expression",
       y = "Distance",
       title = paste0("Monocyte to Kupffer cell: scTenifoldNet")) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_size_continuous(name = "P-value adjusted", range = c(4, 0.1), breaks = c(0.01, 0.05, 0.10, 0.25, 0.5, 1.0)) +
  scale_fill_manual(values = c("deepskyblue3" = "deepskyblue3", "red" = "red")) +
  coord_cartesian(ylim = c(min(diffReg.MonovsKC$distance), max(diffReg.MonovsKC$distance)))

plot_scTFNet

```



```{r}
par(mfrow=c(2,1), mar=c(3,3,1,1), mgp=c(1.5,0.5,0))
set.seed(2711)
qChisq <- rchisq(100,1)
geneColor <- rev(ifelse(scTFNet.MonovsKC$diffRegulation$p.adj < 0.1, 10,1))
qqplot(qChisq, scTFNet.MonovsKC$diffRegulation$FC, pch = 16, main = 'scTFNet.MonovsKC', col = geneColor, 
       xlab = expression(X^2~Quantiles), ylab = 'FC', xlim=c(0,8), ylim=c(0,13))
qqline(qChisq)
legend('bottomright', legend = c('FDR < 0.1'), pch = 16, col = 'red', bty='n', cex = 0.7)
```

# Downstream Anaysis - Joel Pablos Martín
```{r}
TFananse <- read.table("../Results/ANANSE_view/KC/TFactivity_binding.txt", sep="\t", header = TRUE)
TFananse <- TFananse[order(-TFananse$activity),]
TFananse <- TFananse[TFananse$factor != "Alb", ]
rownames(TFananse) <- seq(1,nrow(TFananse))
hist(TFananse$activity, breaks = 50) # the highest peak is at 0.8

# We select the TFs from the diffReg list from scTenifoldNet between Monocyte and KC
diffReg.scTenifoldNet <- diffReg.MonovsKC[diffReg.MonovsKC$gene %in% TFananse$factor,]
plot(diffReg.MonovsKC$distance, diffReg.MonovsKC$FC)
plot(diffReg.MonovsKC$distance, diffReg.MonovsKC$p.adj)

# We select the genes significntly differentially regulated (p-adj < 0.05)
diffReg.MonovsKC.padj <- diffReg.MonovsKC[diffReg.MonovsKC$p.adj <= 0.05,]
diffReg.MonovsKC.padj[,c(1,2)]
```





# scTenifoldNet TUTORIAL
## 1. Simulating of a dataset
Here we simulate a dataset of 2000 cells (columns) and 100 genes (rows) following the negative binomial distribution with high sparsity (~67%). We label the last ten genes as mitochondrial genes ('mt-') to perform single-cell quality control.
```{r}
nCells = 2000
nGenes = 100
set.seed(1)
X <- rnbinom(n = nGenes * nCells, size = 20, prob = 0.98)
X <- round(X)
X <- matrix(X, ncol = nCells)
rownames(X) <- c(paste0('ng', 1:90), paste0('mt-', 1:10))
```

## 2. Generating a perturbed network
```{r}
Y <- X
Y[10,] <- Y[50,]
Y[2,] <- Y[11,]
Y[3,] <- Y[5,]
```

## 3. scTenifoldNet
Here we run scTenifoldNet under the H0 (there is no change in the regulation of the gene) using the same matrix as input and under the HA (there is a change in the regulation of the genes) using the control and the perturbed network.
```{r}
outputH0 <- scTenifoldNet(X = X, Y = X,
                          nc_nNet = 10, nc_nCells = 500,
                          td_K = 3, qc_minLibSize = 30)

outputHA <- scTenifoldNet(X = X, Y = Y,
                          nc_nNet = 10, nc_nCells = 500,
                          td_K = 3, qc_minLibSize = 30)
```

The output of sctTenifoldNet is a list with 3 slots containing: tensorNetworks: The computed weight-averaged denoised gene regulatory networks after CANDECOMP/PARAFAC (CP) tensor decomposition, manifoldAlignment: The generated low-dimensional features result of the non-linear manifold alignment, and diffRegulation: The results of the differential regulation analysis.

## 4. Accessing the computed weight-averaged denoised gene regulatory networks
Networks are provided as matrices of class dgCMatrix that can be easily converted into an igraph object as follows:
```{r}
# Network for sample X
X_igraph <- igraph::graph_from_adjacency_matrix(adjmatrix = outputH0$tensorNetworks$X, weighted = TRUE)
X_igraph
plot(X_igraph)
```

```{r}
# Network for sample Y
Y_igraph <- igraph::graph_from_adjacency_matrix(outputH0$tensorNetworks$Y,weighted = TRUE)
Y_igraph
plot(Y_igraph)
```
## 5. Accessing the manifold alignment result
The generated low-dimensional features result of the non-linear manifold alignment is reported as an object of class data.frame with 2 times the number of tested genes and d dimensions (by default= 30). The information of the X sample is located at the top of the list followed by the one for the Y sample.
```{r}
head(outputH0$manifoldAlignment)
```

For each gene, there are two rows in the manifold alignment, one for each sample:
```{r}
outputH0$manifoldAlignment[c('X_ng59', 'y_ng59'),]
```

```{r}
outputHA$manifoldAlignment[c('X_ng59', 'y_ng59'),]
```


The Euclidean distance is computed for each pair of coordinates (representing a gene in each sample) and all the distances together are used to perform the differential regulation test.
```{r}
dist(outputH0$manifoldAlignment[c('X_ng59', 'y_ng59'),])
dist(outputHA$manifoldAlignment[c('X_ng59', 'y_ng59'),])
```

## 6. Differential regulation based on manifold alignment distances
As is shown below, under the H0, none of the genes shown a significative difference in regulatory profiles using an FDR cut-off of 0.1, but under the HA, the 6 genes involved in the perturbation (50, 11, 2, 10, 5, and 3) are identified as perturbed.
```{r}
head(outputH0$diffRegulation, n = 10)
```

```{r}
head(outputHA$diffRegulation, n = 10)
```

## 7. Plotting the results
Results can be easily displayed using quantile-quantile plots. Here we labeled in red the identified perturbed genes with FDR < 0.1.
```{r}
par(mfrow=c(2,1), mar=c(3,3,1,1), mgp=c(1.5,0.5,0))
set.seed(1)
qChisq <- rchisq(100,1)
geneColor <- rev(ifelse(outputH0$diffRegulation$p.adj < 0.1, 10,1))
qqplot(qChisq, outputH0$diffRegulation$FC, pch = 16, main = 'H0', col = geneColor, 
       xlab = expression(X^2~Quantiles), ylab = 'FC', xlim=c(0,8), ylim=c(0,13))
qqline(qChisq)
legend('bottomright', legend = c('FDR < 0.1'), pch = 16, col = 'red', bty='n', cex = 0.7)

geneColor <- rev(ifelse(outputHA$diffRegulation$p.adj < 0.1, 'red', 'black'))
qqplot(qChisq, outputHA$diffRegulation$FC, pch = 16, main = 'HA', col = geneColor, 
       xlab = expression(X^2~Quantiles), ylab = 'FC', xlim=c(0,8), ylim=c(0,13))
qqline(qChisq)
legend('bottomright', legend = c('FDR < 0.1'), pch = 16, col = 'red', bty='n', cex = 0.7)
```



# IReNA Tutorial: Integrate bulk ATAC-seq and scRNA-seq data to reconstruct regulatory networks
https://jiang-junyao.github.io/IReNA/bulk-ATAC+scRNA#integrate-bulk-atac-seq-and-scrna-seq-data-to-reconstruct-regulatory-networks
```{r}
###call Mus musculus motif database
TRANSFAC18 <- Tranfac201803_Mm_MotifTFsF
TRANSFAC18
```

## Step 1: Calculate the pseudotime
```{r}
### Read seurat_object
library(IReNA)
### Calculate the pseudotime and return monocle object
monocle_object <- get_pseudotime(All_Seurat,gene.use = rownames(All_Seurat))
###Add pseudotime to the Seurat object
### This function only support monocle object from monocle2
All_Seurat_time <- add_pseudotime(All_Seurat, monocle_object)
```






























