---
title: "DiffBind_Lasso_mewtwo"
author: "Joel Pablos Martín"
date: "2023-06-10"
output: html_document
---

# Setting working directory
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir ="/srv/data/joelpm/Code")
setwd("/srv/data/joelpm/Code")
getwd()
```

```{r, eval=FALSE}
# To convert the .rmd to .R
knitr::purl("/srv/data/joelpm/Code/DiffBind_Lasso_mewtwo.rmd", documentation=2)
```

# Load packages
```{r, eval = FALSE}
# install.packages("devtools")
# devtools::install_github("GreenleafLab/chromVARmotifs")
# BiocManager::install("GenomicRanges")
# BiocManager::install("monaLisa")
# BiocManager::install("JASPAR2020")
# devtools::install_github("GreenleafLab/chromVARmotifs")
# BiocManager::install("TFBSTools")
# BiocManager::install("DiffBind")
# BiocManager::install("DirichletMultinomial")
# BiocManager::install("BSgenome.Mmusculus.UCSC.mm39")
# BiocManager::install("SummarizedExperiment")
# BiocManager::install("Biostrings")
# BiocManager::install("ComplexHeatmap")
# BiocManager::install("HDF5Array")
# BiocManager::install("BiocParallel")
# install.packages("snow")
# install.packages("circlize")
# install.packages("Seurat")
```

```{r, results = FALSE, echo = FALSE, message = FALSE}
library(devtools)
library(GenomicRanges)
library(monaLisa)
library(JASPAR2020)
library(TFBSTools)
library(DiffBind)
library(DirichletMultinomial)
library(BSgenome.Mmusculus.UCSC.mm39)
library(SummarizedExperiment)
library(Biostrings)
library(ComplexHeatmap)
library(HDF5Array)
library(BiocParallel)
library(snow)
library(circlize)
library(Seurat)
library(chromVARmotifs)
```



# DiffBind Analysis
## Loading data and experimental design
```{r}
library(DiffBind)
db.NICHE <- read.table('../Data/DiffBind/samplesheet_thesis.tsv',sep = "\t", header=T)
dbObj.NICHE <- dba(sampleSheet=db.NICHE)
dbObj.NICHE
```

## Affinity binding matrix
```{r}
dbObj.NICHE <- dba.count(dbObj.NICHE, bUseSummarizeOverlaps=TRUE) # bParallel =  FALSE
dbObj.NICHE
```

## PCA plot
```{r}
dba.plotPCA(dbObj.NICHE, attributes = DBA_FACTOR)
plot(dbObj.NICHE)
```

## Differential enrichment analysis
```{r}
# Design of the pairwise comparisons
dbObj.NICHE <- dba.contrast(dbObj.NICHE, categories=DBA_FACTOR, minMembers = 2)
# Differential enrichment analysis
dbObj.NICHE <- dba.analyze(dbObj.NICHE, method=dbObj.NICHE$config$AnalysisMethod, bParallel=TRUE)
dbObj.NICHE <- dba.analyze(dbObj.NICHE, method=DBA_DESEQ2, bParallel=TRUE)
dbObj.NICHE <- dba.analyze(dbObj.NICHE, method=DBA_EDGER, bParallel=TRUE)
# save(dbObj.NICHE, ascii=FALSE, file="../Results/DiffBind_Lasso/dbObj_NICHE.bin")
# load("../Results/DiffBind_Lasso/dbObj_NICHE.bin")
```


```{r}
designDB_NICHE <- dba.show(dbObj.NICHE, bContrasts=T)
designDB_NICHE
# dba.plotVenn(dbObj.NICHE, contrast=1,method=DBA_ALL_METHODS)
```

```{r}
dba.plotMA(dbObj.NICHE, contrast = c(2,4,13,21), method=DBA_DESEQ2)
```

```{r}
dba.plotMA(dbObj.NICHE, contrast = c(2,4,13,21), bXY=TRUE)
pvals <- dba.plotBox(dbObj.NICHE, contrast = 21)
```

## Extracting results
```{r}
res_deseq_NICHE <- dba.report(dbObj.NICHE, method=DBA_DESEQ2, contrast = 21, th=1)
res_deseq_NICHE
```

```{r}
setClass("MyClass_NICHE", slots = c(KC = "list", HSC = "list", LSEC = "list", Hep = "list", Neutro = "list", KC_KG = "list", Mono_KG = "list"))
NICHE_DiffBind <- new("MyClass_NICHE", KC = list(), HSC = list(), LSEC = list(), Hep = list(), Neutro = list(), KC_KG = list(), Mono_KG = list())

cell_types <- c("KC","HSC","LSEC","Hep","Neutro","KC_KG","Mono_KG")

for (i in seq(1,nrow(designDB_NICHE))){
  res_deseq_NICHE <- dba.report(dbObj.NICHE, method=DBA_DESEQ2, contrast = i, th=1)
  slot(NICHE_DiffBind, designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_GR")]] <- res_deseq_NICHE
  slot(NICHE_DiffBind, designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_seq")]] <- getSeq(BSgenome.Mmusculus.UCSC.mm39, res_deseq_NICHE)
  rm(res_deseq_NICHE)
}

# save(NICHE_DiffBind, ascii=FALSE, file="../Results/DiffBind_Lasso/dbObjResults_NICHE.bin")
# load("../Results/DiffBind_Lasso/dbObjResults_NICHE.bin")
```


## Prepare TFBM database
```{r, eval = FALSE, echo = FALSE}
# get PFMs (vertebrate TFs from Jaspar)
pfms <- getMatrixSet(JASPAR2020, list(matrixtype = "PFM", 
                                      tax_group = "vertebrates"))

# randomly sample 300 PFMs for illustration purposes (for quick runtime)
set.seed(4563)
pfms <- pfms[sample(length(pfms), size = 300)]

# convert PFMs to PWMs
pwms <- toPWM(pfms)
```

```{r}
# we get the PWMs from Mus musculus TFs from JASPAR database
# PWMsDB <- getMatrixSet(JASPAR2020, opts = list(matrixtype = "PWM", tax_group = "vertebrates", familly = "mus.musculus"))
# length(PWMsDB)
PFMsDB <- readRDS("../Data/GimmeMotifs/CisBP_forR/cisBP_mouse_pfms_2021.rds?download=1")
PWMsDB <- toPWM(PFMsDB) # Convert PFM matrix list to a PWM matrix list with TFBStools
length(PWMsDB)
namesTFs <- names(PWMsDB@listData)
c("Nr1h3", "Rbpj","Rxra","Irf8","Spi1","Spic","Stat1","Stat2") %in% namesTFs
```


## Full similarity matrix
```{r}
# TFBSmatrixCorFull <- cor(TFBSmatrix[, colnames(se)], method = "pearson")
# TFBSmatrixCorFull[1:5,1:5]
suppressWarnings(
PFMmatrixSimFull <-  motifSimilarity(PFMsDB,
                                      y = NULL,
                                      method = "R",
                                      BPPARAM = BiocParallel::SnowParam(workers = 20),
                                      verbose = TRUE)
)
PFMmatrixSimFull[1:5,1:5]
#PFMmatrixSimFull_SRscore <- read.table("../Data/GimmeMotifs/CisBP_forR/Escores.txt", sep = "\t", header = TRUE)

# save(PFMmatrixSimFull, ascii=FALSE, file="../Results/DiffBind_Lasso/PFMmatrixSimFull_CisBP21.bin")
# load("../Results/DiffBind_Lasso/PFMmatrixSimFull_CisBP21.bin")
```

We scan for the motifs from the TFs in the differentially ATAC-seq peaks
```{r}
# get TFBS on given GRanges (peaks)
# suppress warnings generated by matchPWM due to the presence of Ns 
# in the sequences
library(BiocParallel)
library(parallel)
# detectCores()

for (i in seq(1, nrow(designDB_NICHE))){
#for (i in seq(20, nrow(designDB_NICHE))){
  
  peakSeq <- slot(NICHE_DiffBind,designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_seq")]]
  
  set.seed(2711)
  suppressWarnings({
    hits <- findMotifHits(query = PWMsDB, subject = peakSeq, min.score = 10.0,
                          BPPARAM = BiocParallel::MulticoreParam(30))
  }) # BiocParallel::SnowParam(workers = 16)
  
  # get TFBS matrix
  TFBSmatrix <- unclass(table(factor(seqnames(hits), levels = seqlevels(hits)),
                              factor(hits$pwmname, levels = name(PWMsDB))))
  TFBSmatrix[1:6, 1:6]
  
  # remove TF motifs with 0 binding sites in all regions
  zero_TF <- colSums(TFBSmatrix) == 0
  sum(zero_TF)
  #> [1] 2
  TFBSmatrix <- TFBSmatrix[, !zero_TF]
  
  # calculate G+C and CpG obs/expected
  fMono <- oligonucleotideFrequency(peakSeq, width = 1L, as.prob = TRUE)
  fDi <- oligonucleotideFrequency(peakSeq, width = 2L, as.prob = TRUE)
  fracGC <- fMono[, "G"] + fMono[, "C"]
  oeCpG <- (fDi[, "CG"] + 0.01) / (fMono[, "G"] * fMono[, "C"] + 0.01)
  
  # add GC and oeCpG to predictor matrix
  TFBSmatrix <- cbind(fracGC, oeCpG, TFBSmatrix)
  TFBSmatrix[1:6, 1:6]
  
  # We create the SE object and calculate the randomized lasso stability selection
  rownames(TFBSmatrix) <- seq(1,nrow(TFBSmatrix))
  TFBSmatrix <- as.matrix(TFBSmatrix)
  slot(NICHE_DiffBind, designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_TFBSmatrix")]] <- TFBSmatrix
  
  res_deseq_NICHE <- slot(NICHE_DiffBind,designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_GR")]]

  set.seed(2711)
  se <- randLassoStabSel(x = TFBSmatrix[,-c(1,2)], y = res_deseq_NICHE$Fold, 
                         cutoff = 0.8)
  
  slot(NICHE_DiffBind,designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_LassoSE")]] <- se

  # save(NICHE_DiffBind, ascii=FALSE, file="../Results/DiffBind_Lasso/dbObjResults_withLASSO_NICHE_ALREADYSAVED.bin")
  # load("../Results/DiffBind_Lasso/dbObjResults_withLASSO_NICHE.bin")
}

suppressWarnings(rm(se, pvals, peakSeq, fDi, fMono, i, oeCpG, fracGC, zero_TF, namesTFs, hits, res_deseq_NICHE))
```



```{r}
directional = TRUE
selProbMin = 0.8
selProbMinPlot = 0.4
showSelProbMin = TRUE
col = c("cadetblue", "grey", "red")
method = "pearson" # c("pearson", "kendall", "spearman")
ylimext = 0.25
legend = "topright"
legend.cex = 1.0

plotSelectionProb <- function(se,
                              directional = TRUE,
                              selProbMin = metadata(se)$stabsel.params.cutoff, 
                              selProbMinPlot = 0.4,
                              showSelProbMin = TRUE,
                              col = c("cadetblue", "grey", "red"),
                              method = c("pearson", "kendall", "spearman"),
                              ylimext = 0.25,
                              legend = "topright", 
                              legend.cex = 1.0, 
                              cex.main = 0.6,
                              title = "",
                              ...) {
    
    # checks
    # .assertScalar(x = directional, type = "logical")
    # .assertScalar(x = selProbMin, type = "numeric", rngIncl = c(0, 1))
    # .assertScalar(x = selProbMinPlot, type = "numeric", rngIncl = c(0, 1))
    # .assertScalar(x = showSelProbMin, type = "logical")
    # .assertScalar(x = legend, type = "character")
    # .assertScalar(x = legend.cex, type = "numeric", rngExcl = c(0, Inf))
    stopifnot(exprs = {
        is(se, "SummarizedExperiment")
        selProbMin >= selProbMinPlot
    })
    # .assertVector(x = col, len = 3L)
    method <- match.arg(method)
    # .assertScalar(x = ylimext, type = "numeric", rngIncl = c(0, Inf))

    # selection probabilities * sign(correlation to y)
    probs <- se$selProb
    cols <- ifelse(probs > selProbMin, col[1], col[2])
    if (directional) {
        corcoef <- as.vector(cor(x = SummarizedExperiment::rowData(se)$y,
                                 y = SummarizedExperiment::assay(se, "x"),
                                 method = method))
        probs <- probs * sign(corcoef)
    }

    # kept and ordered
    keep <- which(abs(probs) >= selProbMinPlot)
    keep <- keep[order(probs[keep], decreasing = TRUE)]
    cols <- cols[keep]
    predNames <- colnames(se)[keep]
    probs <- probs[keep]
    up <- probs > 0
    selection <- data.frame(predNames = predNames, probs = probs)
    selection <- selection[which(abs(selection$probs) >= selProbMin),]
    rownames(selection) <- seq(1,nrow(selection))
    

    # plot
    if (any(keep)) {
        ret <- graphics::barplot(probs, col = cols, border = NA,
                                        ylab = ifelse(
                                            directional, 
                                            "Directional selection probability",
                                            "Selection probability"
                                        ),
                                        names.arg = NA, axes = FALSE,
                                        ylim = c(min(probs) - ylimext,
                                                 max(probs) + ylimext),
                                        main = title,
                                        ...)
        ys <- pretty(x = c(0, probs))
        graphics::axis(side = 2, at = ys)
        if (showSelProbMin) {
            hval <- if (directional) c(-1, 1) * selProbMin else selProbMin
            graphics::abline(h = hval, lty = 5, col = col[3])
        }
        graphics::legend(x = legend, bty = "n", fill = col[seq_len(2)], 
                         border = NA, legend = c("selected", "not selected"), 
                         cex = legend.cex)
        if (any(up)) {
            graphics::text(x = ret[up], y = probs[up] + par("cxy")[2] / 3,
                           labels = predNames[up], col = cols[up],
                           xpd = TRUE, srt = 90, adj = c(0, 0.5))
        }
        if (any(!up)) {
            graphics::text(x = ret[!up], y = probs[!up] - par("cxy")[2] / 3,
                           labels = predNames[!up], col = cols[!up],
                           xpd = TRUE, srt = 90, adj = c(1, 0.5))
        }
    } else{
        ret <- NULL
    }
    
    list(plot = invisible(ret), probs = selection)
}
suppressWarnings(rm(directional, selProbMin, selProbMinPlot, showSelProbMin, col, method, ylimext, legend, legend.cex, keep, up, probs, cols))
```

The selected TFs are enlarged to include other TFs with a similar motif, and the PFM similarity in the resulting matrix is calculated.
```{r}
setClass("MyClass_NICHE", slots = c(KC = "list", HSC = "list", LSEC = "list", Hep = "list", Neutro = "list", KC_KG = "list", Mono_KG = "list"))
NICHE_DiffBind_expr <- new("MyClass_NICHE", KC = list(), HSC = list(), LSEC = list(), Hep = list(), Neutro = list(), KC_KG = list(), Mono_KG = list())

cell_types <- c("KC","HSC","LSEC","Hep","Neutro","KC_KG","Mono_KG")

for (i in seq(1, nrow(designDB_NICHE))){

  se <- slot(NICHE_DiffBind,designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i],"_LassoSE")]]
  selection <- plotSelectionProb(se, directional = TRUE, title = paste0(designDB_NICHE$Group2[i], " (-) vs ", designDB_NICHE$Group[i], " (+)"))
  selection <- selection[["probs"]]
  df <- data.frame(PFMmatrixSimFull[selection$predNames,])
  
  tf_expanded <- c()
  selection$expansion <- rep(NA,nrow(selection))
  for (tf in seq(1,nrow(df))){
    TF <- selection$predNames[tf]
    tf_list <- unlist(apply(df[tf, -1], 1, function(row) {
                                        col_names <- colnames(df)[-1]
                                        high_indices <- which(row > 0.8)
                                        high_family <- col_names[high_indices]
                                        if (length(high_family) > 0) {
                                          return(high_family)
                                        } else {
                                          return(NA)
                                        }
                                      }))

    selection$expansion[tf] <-  paste(tf_list[-1], collapse = ",")
    p <- as.numeric(selection$probs[tf])
    for (x in tf_list[-1]){
      if (x != TF){
        P <- p * df[tf,x]
        tobind <- c(x, P, TF)
        if (x %in% selection$predNames){
          n <- which(selection$predNames == x)
          N <- c(as.numeric(selection$probs[n]), P)
          selection$probs[n] <- N[which.max(abs(N))]
        }else{
          selection <- rbind(selection, tobind)
        }
      }
    }
  }
  selection$probs  <- round(as.numeric(selection$probs), 2)
  selection <- selection[order(-selection$probs),]
  rownames(selection) <- seq(1,nrow(selection))
  
  # Positive values
  slot(NICHE_DiffBind_expr, designDB_NICHE$Group[i])[[paste0(designDB_NICHE$Group[i],"vs",designDB_NICHE$Group2[i])]] <- selection[selection$probs > 0,]
  # Negative values
  selectionMin <- selection[selection$probs < 0,]
  selectionMin$probs <- abs(selectionMin$probs)
  selectionMin <- selectionMin[order(-selectionMin$probs),]
  rownames(selectionMin) <- seq(1,nrow(selectionMin))
  slot(NICHE_DiffBind_expr, designDB_NICHE$Group2[i])[[paste0(designDB_NICHE$Group2[i],"vs",designDB_NICHE$Group[i])]] <- selectionMin
  
  
}
# save(NICHE_DiffBind_expr, ascii=FALSE, file="../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_raw.bin")
# load("../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_raw.bin")
suppressWarnings(rm(se, selection, df, tf_expanded, TF, tf_list, p, x, P, tobind, n, N, selectionMin, i, tf))
```

## Expression filtering
```{r}
# load("../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_raw.bin")

exprData <- read.table("../Data/ATLAS/MouseStSt/All_pseudoBulkCell.tsv", sep = "\t", header = TRUE, row.names = 1)
exprData
CELLS =  c("KC","HSC","LSEC","Hep","Neutro","KC_KG","Mono_KG")

for (cell in CELLS){
  slts <- slot(NICHE_DiffBind_expr, cell)
  snam <- names(slts)
  take <- match(CELLS, gsub(paste0(cell,"vs"), "", snam))
  take <- take[!is.na(take)]
  slts <- slts[take]
  snam <- names(slts)
  # load expression data
  expCell <- exprData[,grep(gsub("_KG", "",cell),colnames(exprData))]

  for (x in seq(1,length(slts))){
    cell1vs2_KG <- strsplit(snam[x],"vs")[[1]]
    cell1vs2 <- gsub("_KG", "", cell1vs2_KG)
    if (cell1vs2[1] != cell1vs2[2]){
      path_DEG = paste0("../Data/ATLAS/MouseStSt/specificScore/PRESTO/VS_DEGs/PRESTO_DEGs_",cell1vs2[2],"vs",cell1vs2[1],".tsv")
      DEG <- read.table(path_DEG, sep = "\t", header = TRUE, row.names = 1)
      
      DFB <- slts[[x]]
      DFBexp <- merge(DFB, expCell, by.x = "predNames", by.y = "row.names", all.x = TRUE)
      DFBexp <- DFBexp[complete.cases(DFBexp),]
      colnames(DFBexp) <- gsub(paste0(cell1vs2[1],"_"), "", colnames(DFBexp))
      
      # We filter out genes that are expressed in less than 5% of the cells
      DFBexp <- DFBexp[DFBexp$PercCellsExpr > 5.0 ,]
      DFBexp <- DFBexp[order(-DFBexp$probs, -DFBexp$PercCellsExpr),]
      
      # We add the log2FC data
      DEG$Gene <- rownames(DEG)
      DEG <- DEG[,c("Gene","log2FoldChange")]
      DFBexp_mg <- merge(DFBexp, DEG, by.x = "predNames", by.y = "Gene", all.x = TRUE)
      colnames(DFBexp_mg)[ncol(DFBexp_mg)] <- paste0("log2FC_",gsub("_KG", "", snam[x]))
      DFBexp_mg[, ncol(DFBexp_mg)] <- ifelse(is.na(DFBexp_mg[, ncol(DFBexp_mg)]), 0, DFBexp_mg[, ncol(DFBexp_mg)])
      DFBexp_mg <- DFBexp_mg[order(-DFBexp_mg$probs, -DFBexp_mg[,ncol(DFBexp_mg)]),]
      rownames(DFBexp_mg) <- seq(1,nrow(DFBexp_mg))

      # We save the results
      slot(NICHE_DiffBind_expr, cell)[[snam[[x]]]] <- DFBexp_mg
    }
  }
}
suppressWarnings(rm(slts, snam, take, expCell, cell1vs2_KG, cell1vs2, path_DEG, DEG, DFB, DFBexp, DFBexp_mg, x, cell))
# save(NICHE_DiffBind_expr, ascii=FALSE, file="../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_filt.bin")
# load("../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_filt.bin")
```


Gene expression filtering is applied for the expanded TF list
```{r}
load("../Results/DiffBind_Lasso/dbObjResults_NICHE_DiffBind_expr_filt.bin")

CELLS1 <- c("KC","HSC","LSEC","Hep","Neutro","Mono_KG")
CELLS2 <- c("KC","HSC","LSEC","Hep","Neutro")
CELLS3 <- c("KC","HSC","LSEC","Hep")
CELLS4 <- c("KC_KG","Mono_KG")
CELLS_LIST <- list(CELLS1, CELLS2, CELLS3, CELLS4)
#CELLS_LIST <- list(CELLS3)

m = 0
for (CELLS in CELLS_LIST){
  m = m + 1
  colCRM <- CELLS
  for (cell in CELLS){
    slts <- slot(NICHE_DiffBind_expr, cell)
    snam <- names(slts)
    take <- match(CELLS, gsub(paste0(cell,"vs"), "", snam))
    take <- take[!is.na(take)]
    slts <- slts[take]
    snam <- names(slts)
    
    colCRM <- snam
    zerorw <- rep(c(0), each = length(colCRM))
    CRM <- data.frame(matrix(ncol = length(colCRM), nrow = 0))
    colnames(CRM) <- colCRM
    # Load differential expression the data of the cell type
    cellnam <- gsub("_KG","",cell)
    expCell <- exprData[,grep(cellnam,colnames(exprData))]

    for (x in seq(1,length(slts))){
      vsCell = snam[x]
      for (r in slts[[x]]$predNames){
        if (r %in% rownames(CRM)){
          CRM[r,vsCell] <- 1
        }else{
            CRM[r,] <- zerorw
            CRM[r,vsCell] <- 1
            # CRM[r,c("tpm","PercCellsExpr")] <- expCell[r,]
           }
        }
      }
    
    # Now we calculate in how many cell types does each TF is present
    message(CELLS)
    CRM$TotalSpec <- rowSums(CRM)
    CRM <- CRM[order(-CRM$TotalSpec, rownames(CRM)),]
    slot(NICHE_DiffBind_expr, cell)[[paste0("COMMON",m,"_",cell)]] <- CRM
    # CRM <- NICHE_DiffBind_expr@KC[["COMMON3_KC"]]
    
    # We merge the comparisons by doing the mean
    otherCells <- CELLS[!CELLS %in% cell]
    
    slts <- slot(NICHE_DiffBind_expr, cell)
    snam <- names(slts)
    take <- match(otherCells, gsub(paste0(cell,"vs"), "", snam))
    take <- take[!is.na(take)]
    slts <- slts[take]
    snam <- names(slts)
    
    comp_list <- list()
    for (other in seq(1,length(snam))){
      # DFBLS <- subset(slts[[other]], select = c("predNames","expansion","probs"))
      DFBLS <- slts[[other]][,c(1,2,4,5,6)]
      colOrignal <- colnames(DFBLS)
      colnames(DFBLS) <- paste0(other,".", colnames(DFBLS))
      colnames(DFBLS)[c(1,3,4)] <- c("predNames","tpm","PercCellsExpr")
      comp_list[[snam[[other]]]] <- DFBLS
      
      if (other == 1){First <- DFBLS}
      else{
        DFBLS_merge <- merge(First, DFBLS, by = "predNames", all = TRUE)
        DFBLS_merge[is.na(DFBLS_merge)] <- 0
        First <- DFBLS_merge
        }
      if (length(snam) == 1){
        DFBLS_merge <- First
        DFBLS_merge[is.na(DFBLS_merge)] <- 0
        }
    }
    
    DFBLS_merge <- DFBLS_merge[, -grep("tpm", colnames(DFBLS_merge))]
    DFBLS_merge <- DFBLS_merge[, -grep("PercCellsExpr", colnames(DFBLS_merge))]
    DFBLS_final <- merge(DFBLS_merge, expCell, by.x = "predNames", by.y = "row.names", all.x = TRUE)
    colnames(DFBLS_final) <- gsub(paste0(cellnam,"_"),"", colnames(DFBLS_final))
    # DFBLS_merge <- DFBLS_merge[, c("predNames", "expansion", setdiff(colnames(DFBLS_merge), c("predNames", "expansion")))]
    head(DFBLS_merge)
    # DFBLS_final <- data.frame(factor = DFBLS_merge$predNames, row.names = seq(1,nrow(DFBLS_merge)))
    
    # Now we do the mean of each value in the dataframe
    if (length(snam) != 1){
      DFBLS_final$probs <- rowMeans(DFBLS_final[,grep("\\.probs", colnames(DFBLS_final))])
      DFBLS_final$log2FC <- rowMeans(DFBLS_merge[,grep("\\.log2FC", colnames(DFBLS_final))])
    }else{
      DFBLS_final$probs <- DFBLS_final[,grep("\\.probs", colnames(DFBLS_final))]
      DFBLS_final$log2FC <- DFBLS_merge[,grep("\\.log2FC", colnames(DFBLS_final))]
    }
    DFBLS_final <- DFBLS_final[, -grep("\\.probs", colnames(DFBLS_final))]
    DFBLS_final <- DFBLS_final[, -grep("\\.log2FC", colnames(DFBLS_final))]
    
    # We merge them with he specificity score
    DFBLS_final_merge <- merge(DFBLS_final, CRM, by.x = "predNames", by.y = "row.names", all = TRUE)
    DFBLS_final_merge$TotalSpec[is.na(DFBLS_final_merge$TotalSpec)] <- 0
    DFBLS_final_merge$NewProb <- (DFBLS_final_merge$probs * DFBLS_final_merge$TotalSpec + 1)/length(CELLS)
    DFBLS_final_merge <- DFBLS_final_merge[order(-DFBLS_final_merge$NewProb),]
    rownames(DFBLS_final_merge) <- seq(1,nrow(DFBLS_final_merge))
    slot(NICHE_DiffBind_expr, cell)[[paste0("COMMON",m,"_",cell)]] <- DFBLS_final_merge
    
    # We filter out TFs thattttt are more expressed in the other cell types than the mainCell type
    DFBLS_final_merge <- DFBLS_final_merge[DFBLS_final_merge$log2FC > 0,]

    # Create the scatter plot with modified aesthetics and non-overlapping labels
    thresh = 25
    top_labels1 <- DFBLS_final_merge[order(DFBLS_final_merge$probs, decreasing = TRUE), ][1:thresh, ]
    plot1 <- ggplot(data = DFBLS_final_merge, aes(x = log2FC, y = probs)) +
            geom_point(aes(size = PercCellsExpr) , shape = 21, fill = "deepskyblue3", color = "deepskyblue4", alpha = 0.4) +
            geom_text_repel(data = subset(top_labels1, probs > 0.0),
                            aes(label = predNames), box.padding = 0.5,  max.overlaps = Inf,  force = 0.5) +
            labs(x = "Log2 fold change of TF",
                 y = "Probability",
                 title = paste0(cell," - DiffBind + Lasso.Rand.Stab.Sel. (COMMON ",m,")")) +
            theme_classic() +
            theme(plot.title = element_text(hjust = 0.5)) + # face = "bold",family = "Times New Roman"
            scale_size_continuous(name = "% cells with expression", 
                                  breaks = c(0, 25, 50, 75, 100),
                                  labels = c(0, 25, 50, 75, 100),
                                  limits = c(0, 100)) 
            # + coord_cartesian(ylim = c(min(DFBLS_final_merge$probs), 1.25))
    
    top_labels2 <- DFBLS_final_merge[order(DFBLS_final_merge$NewProb, decreasing = TRUE), ][1:thresh, ]
    plot2 <- ggplot(data = DFBLS_final_merge, aes(x = log2FC, y = NewProb)) +
        geom_point(aes(size = PercCellsExpr) , shape = 21, fill = "deepskyblue3", color = "deepskyblue4", alpha = 0.4) +
        geom_text_repel(data = subset(top_labels2, NewProb > 0.0),
                        aes(label = predNames), box.padding = 0.5, max.overlaps = Inf,  force = 0.5) +
        labs(x = "Log2 fold change of TF",
             y = "Probability",
             title = paste0(cell," - DiffBind + Lasso.Rand.Stab.Sel. (COMMON ",m,")")) +
        theme_classic() +
        theme(plot.title = element_text(hjust = 0.5)) + # face = "bold",family = "Times New Roman"
        scale_size_continuous(name = "% cells with expression", 
                                  breaks = c(0, 25, 50, 75, 100),
                                  labels = c(0, 25, 50, 75, 100),
                                  limits = c(0, 100)) 
        # + coord_cartesian(ylim = c(min(DFBLS_final_merge$NewProb), 1.25))
    
    # Save the plot as a high-resolution JPEG
    pathjpg1 = paste0("../Results/DiffBind_Lasso/plots/",cell,"/DiffBind_",cell,"_",m,"_probMeans.jpg")
    ggsave(pathjpg1, plot1, dpi = 300, width = 8, height = 6, units = "in")
    pathjpg2 = paste0("../Results/DiffBind_Lasso/plots/",cell,"/DiffBind_",cell,"_",m,"_probMeansANDspec.jpg")
    ggsave(pathjpg2, plot2, dpi = 300, width = 8, height = 6, units = "in")
    
    # Display the plot
    message(paste0(cell," - DiffBind + Lasso.Rand.Stab.Sel. (COMMON ",m,")"))
    print(plot1)
    print(plot2)
  }
  
}
# save(NICHE_DiffBind_expr, ascii=FALSE, file="../Results/DiffBind_Lasso/NICHE_DiffBind_expr_FINAL.bin")
# load("../Results/DiffBind_Lasso/NICHE_DiffBind_expr_FINAL.bin")
```


```{r}
# heatmap
selTFs <- NICHE_DiffBind_expr@Mono_KG[["Mono_KGvsKC_KG"]][["predNames"]]
pfmsSel <- PFMsDB[match(selTFs, name(PFMsDB))]
maxwidth <- max(sapply(TFBSTools::Matrix(pfmsSel), ncol))
seqlogoGrobs <- lapply(pfmsSel, seqLogoGrob, xmax = maxwidth)
se <- NICHE_DiffBind@Mono_KG[["Mono_KGvsKC_KG_LassoSE"]]

hmSeqlogo <- rowAnnotation(logo = annoSeqlogo(seqlogoGrobs, which = "row"),
                           annotation_width = unit(2, "inch"), 
                           show_annotation_name = FALSE
)

colAnn <- HeatmapAnnotation(AUC = se$selAUC, selProb = se$selProb,
                            show_legend = TRUE, 
                            show_annotation_name = TRUE,
                            col = list(
                              AUC = colorRamp2(c(0, 1), 
                                               c("white", "brown")),
                              selProb = colorRamp2(c(0, 1), 
                                                   c("white", "steelblue")))
)

Heatmap(TFBSmatrixCorSel, 
        show_row_names = TRUE, 
        show_column_names = TRUE, 
        name = "Pear. Cor.", column_title = "Selected TFs",
        col = colorRamp2(c(-1, 0, 1), c("blue", "white", "red")), 
        right_annotation = hmSeqlogo,
        top_annotation = colAnn)
```




