---
title: "ANANSE_validation_qPCR_and_cytoscape"
author: "Joel Pablos Martin"
date: "2023-04-07"
output: html_document
---


```{r}
setwd("/srv/data/joelpm/Code")
```

```{r}
setClass("MyClass_NICHE", slots = c(KC = "list", HSC = "list", LSEC = "list", Hep = "list", Neutro = "list"))
NICHE_qPCR <- new("MyClass_NICHE", KC = list(), HSC = list(), LSEC = list(), Hep = list(), Neutro = list())
```

```{r}
#Load the data
#cell_types <- c("KC","HSC","LSEC","Hep","Neutro","Mono_KG","KC_KG")
cell_types <- c("HSC")

dbpx <- "" #dbpx <- "gimmemm39_"
for (cell in cell_types){
  slot(NICHE_qPCR,cell)[["GRN"]] <-   read.table(paste0("../Results/ANANSE_network/",cell,"/",dbpx,"ANANSE_network_CYTOSCAPE.tsv"), header = TRUE, sep = "\t")
}
slot(NICHE_qPCR,cell)[["GRN"]]
```

# qPCR validation : today 2023/05/13
Load specificity scores from the atlas
```{r}
cell = "Stellate_cells"
presto.cell <- read.table(paste0("../Data/ATLAS/MouseStSt/specificScore/PRESTO/PRESTO_specificScore_", cell, ".tsv"), header = TRUE, sep = "\t", row.names = 1)
presto.cell
```


```{r}
cell = "HSC"
GRN <- slot(NICHE_qPCR, cell)[["GRN"]]
allTFs <- unique(GRN$tf)
GRN$is.target.a.TF <- ifelse(GRN$target %in% allTFs, "yes", "no")
```


```{r}
core_node = "Rxra"
GRN_node <- GRN[GRN$tf == core_node,]
GRN_node
```

Calculating the final score by merging the probability and the specificity ranked scores
```{r}
GRN_node <- GRN_node[GRN_node$target %in% presto.cell$feature,]
GRN_node <- GRN_node[order(GRN_node$target),]
presto.cell <- presto.cell[order(presto.cell$feature),]
GRN_node <- cbind(GRN_node, rank.pct_diff = presto.cell$rank.pct_diff, rank.FC = presto.cell$rank.FC, specif.score = presto.cell$specif.score, rank.specific = presto.cell$rank.specific)

# Calculation of the selection score
GRN_node <- GRN_node[order(-GRN_node$prob),]
GRN_node$rank.prob <- seq(nrow(GRN_node),1)
rownames(GRN_node) <- seq(nrow(GRN_node),1)

GRN_node$selection.score <- GRN_node$rank.prob * GRN_node$rank.specific
GRN_node$selection.score <- (GRN_node$selection.score - min(GRN_node$selection.score)) / (max(GRN_node$selection.score) - min(GRN_node$selection.score))
GRN_node <- GRN_node[order(-GRN_node$selection.score),]
GRN_node$rank.selection <- seq(1:nrow(GRN_node))
rownames(GRN_node)<- seq(1:nrow(GRN_node))
GRN_node

write.table(GRN_node, paste0("../Results/ANANSE_network/",cell,"/ANANSE_network_",core_node,"_finalScore.tsv"), sep="\t", row.names=TRUE, col.names = NA, quote=FALSE)
plot(GRN_node$rank.prob, GRN_node$rank.specific)
```

List of TFs and target for qPCR validation
```{r}
GRN_node[1:30,c(1,2,3,8,10,12)]
GRN_node[GRN_node$is.target.a.TF == "no",][1:30,c(1,2,3,8,10,12)]
node_TFs <- GRN_node[GRN_node$is.target.a.TF == "yes",][1:30,c(1,2,3,8,10,12)]
# rownames(node_TFs) <- seq(1,nrow(node_TFs))
node_TFs


node_TFs[node_TFs$target == "Spic",]
GRN_node[GRN_node$target == "Spic",]
GRN_node[GRN_node$target == "Tfec",]

```

```{r}
library(ggplot2)
library(ggrepel)

highlight_genes <- c("Clec4f", "Spi1", "C1qc", "Cd5l", "Spic", "Nr1h3", "Rbpj")
highlight_genes <- unique(c("Clec4f", "Spi1", "C1qc", "Cd5l", "Spic", "Nr1h3", "Rbpj", head(GRN_node$target, 8)))
highlighted_rows <- which(GRN_node$target %in% highlight_genes)

plot_data <- data.frame(x = GRN_node$rank.prob, y = GRN_node$rank.specific, gene = GRN_node$target)

plot <- ggplot(plot_data, aes(x, y)) +
  geom_point(color = "black", size = 0.5) +
  geom_point(data = plot_data[highlighted_rows, ], aes(x, y), color = "red", size = 1) +
  geom_label_repel(data = plot_data[highlighted_rows, ], aes(label = gene), color = "red", nudge_x = 0.2, nudge_y = 0.1, max.overlaps = Inf) +
  xlab("rank.prob") +
  ylab("rank.specific") +
  ggtitle("Kupffer Cells final score ranking") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5), plot.title.position = "plot")

plot
ggsave(paste0("../Results/plots/specific_score/ananse_",core_node,"_finalScore_",cell,".jpg"), plot, dpi = 300)
```



# Past analysis

```{r}
cell_types <- c("KC")
for (cell in cell_types){
  GRN <- slot(NICHE_qPCR,cell)[["GRN"]]
  filtrows <- subset(GRN, prob > 0.60 & tf_expression > 0.75 & target_expression > 0.80 & weighted_binding > 0.50)
  filtrows <- subset(GRN, prob > 0.9 & tf_expression > 0.7 & target_expression > 0.70 & weighted_binding > 0.50)
  message(dim(filtrows)*100/dim(GRN))
  indegree <- table(filtrows$target)
  outdegree <- table(filtrows$tf)
  # transcription factor degrees
  in_tf <- as.vector(indegree[GRN$tf])
  in_tf[is.na(in_tf)] <- 0
  norm_in_tf = round((in_tf - min(in_tf)) / (max(in_tf) - min(in_tf)),4)
  out_tf <- as.vector(outdegree[GRN$tf])
  out_tf[is.na(out_tf)] <- 0
  norm_out_tf = round((out_tf - min(out_tf)) / (max(out_tf) - min(out_tf)),4)
  # target genes degrees
  in_target <- as.vector(indegree[GRN$target])
  in_target[is.na(in_target)] <- 0
  norm_in_target = round((in_target - min(in_target)) / (max(in_target) - min(in_target)),4)
  out_target <- as.vector(outdegree[GRN$target])
  out_target[is.na(out_target)] <- 0
  norm_out_target = round((out_target - min(out_target)) / (max(out_target) - min(out_target)),4)

  message(mean(in_tf),' ',mean(out_tf),' ',mean(in_target),' ',mean(out_target))
  message(mean(norm_in_tf),' ',mean(norm_out_tf),' ',mean(norm_in_target),' ',mean(norm_out_target))
  
  GRN_inout <- cbind(GRN,norm_in_tf,norm_out_tf,norm_in_target,norm_out_target)
  slot(NICHE_qPCR,cell)[["GRN_inout"]] <- GRN_inout
  GRN_inout
  # slot(NICHE_qPCR,cell)[["GRN_filteredges"]] <-  
}
# Rbpj
GRN_tf_Rbpj <- GRN_inout[GRN_inout$tf == 'Rbpj',]
GRN_tf_Rbpj <- GRN_tf_Rbpj[order(GRN_tf_Rbpj$prob, decreasing = T),]
norm_GRN_tf_Rbpj <- apply(GRN_tf_Rbpj[,8:11], 2, function(x) (x - min(x)) / (max(x) - min(x)))
norm_GRN_tf_Rbpj <- cbind(GRN_tf_Rbpj[,1:7],norm_GRN_tf_Rbpj)
norm_GRN_tf_Rbpj
norm_GRN_tf_Rbpj$score <- 0.5 * norm_GRN_tf_Rbpj$prob + 0.5 * (1-norm_GRN_tf_Rbpj$norm_in_target)
norm_GRN_tf_Rbpj <- norm_GRN_tf_Rbpj[order(norm_GRN_tf_Rbpj$score, decreasing = TRUE),]
norm_GRN_tf_Rbpj
```



```{r}
library(tibble)

GRN_tf_Rbpj <- GRN_inout[GRN_inout$tf == 'Rbpj',]
GRN_tf_Rbpj <- GRN_tf_Rbpj[order(GRN_tf_Rbpj$prob, decreasing = TRUE),]

norm_GRN_tf_Rbpj <- GRN_tf_Rbpj[, 8:11]
for (col in names(norm_GRN_tf_Rbpj)) {
  col_range <- max(norm_GRN_tf_Rbpj[[col]]) - min(norm_GRN_tf_Rbpj[[col]])
  if (col_range == 0) {
    norm_GRN_tf_Rbpj[[col]] <- ifelse(is.nan(norm_GRN_tf_Rbpj[[col]]), NaN, 1)
  } else {
    norm_GRN_tf_Rbpj[[col]] <- (norm_GRN_tf_Rbpj[[col]] - min(norm_GRN_tf_Rbpj[[col]])) / col_range
  }
}

norm_GRN_tf_Rbpj <- tibble::rownames_to_column(norm_GRN_tf_Rbpj, var = "target")
norm_GRN_tf_Rbpj <- cbind(GRN_tf_Rbpj[, 1:2], norm_GRN_tf_Rbpj)
norm_GRN_tf_Rbpj
```


```{r}
# Nr1h3
GRN_tf_Nr1h3 <- GRN_inout[GRN_inout$tf == 'Nr1h3',]
GRN_tf_Nr1h3 <- GRN_tf_Nr1h3[order(GRN_tf_Nr1h3$prob, decreasing = T),]
GRN_tf_Nr1h3
which(GRN_tf_Nr1h3$target == KO_Nr1h3)
KO_Nr1h3 <- c('Clec4f','Cx3cr1','Acp5','Cd5l','Kcna2','Il18bp','Arg2')
KGlass_Nr1h3 <- GRN_tf_Nr1h3[GRN_tf_Nr1h3$target %in% KO_Nr1h3,]
KGlass_Nr1h3
which(GRN_tf_Nr1h3$target %in% KO_Nr1h3)

# Normalization Nr1h3
norm_GRN_tf_Nr1h3 <- apply(GRN_tf_Nr1h3[,8:11], 2, function(x) (x - min(x)) / (max(x) - min(x)))
norm_GRN_tf_Nr1h3 <- cbind(GRN_tf_Nr1h3[,1:7],norm_GRN_tf_Nr1h3)
norm_GRN_tf_Nr1h3$score <- 0.5 * norm_GRN_tf_Nr1h3$prob + 0.5 * (1-norm_GRN_tf_Nr1h3$norm_in_target)
norm_GRN_tf_Nr1h3 <- norm_GRN_tf_Nr1h3[order(norm_GRN_tf_Nr1h3$score, decreasing = TRUE),]
norm_GRN_tf_Nr1h3
which(norm_GRN_tf_Nr1h3$target %in% KO_Nr1h3)

# Smad4
GRN_tf_Smad4 <- GRN_inout[GRN_inout$tf == 'Smad4',]
GRN_tf_Smad4 <- GRN_tf_Smad4[order(GRN_tf_Smad4$prob, decreasing = T),]
GRN_tf_Smad4
KO_Smad4 <- c('Clec4f','Nr1h3','Id3','Spic','Cx3cr1')
which(GRN_tf_Smad4$target == KO_Smad4)
KGlass_Smad4 <- GRN_tf_Smad4[GRN_tf_Smad4$target %in% KO_Smad4,]
KGlass_Smad4
which(GRN_tf_Smad4$target %in% KO_Smad4)
```


```{r}
library(gplots)  # Load the gplots library

# Convert the data frame into a matrix
mat <- data.matrix(KGlass_Nr1h3[, 3:7])  # Select columns 3 to 7 as the data for the heatmap
rownames(mat) <- KGlass_Nr1h3$target

mat_Nr1h3 <- data.matrix(GRN_tf_Nr1h3[, 3:7])
rownames(mat_Nr1h3) <- GRN_tf_Nr1h3$target
# Create the heatmap
heatmap.2(mat, dendrogram = "both", Rowv = TRUE, Colv = FALSE, col = heat.colors(50))
heatmap.2(mat_Nr1h3, dendrogram = "both", Rowv = TRUE, Colv = FALSE, col = heat.colors(50))
# Heatmap 
heatmap(GRN_tf_Nr1h3, aes(GRN_tf_Nr1h3$tf, GRN_tf_Nr1h3$target, fill= GRN_tf_Nr1h3$prob)) + 
  geom_tile()

# Library
library(ggplot2)

# Dummy data


```

# Load packages
```{r}
# install.packages("igraph")
# install.packages("leiden")
library(leiden)
library(Seurat)
library(igraph)
```


```{r}
library(igraph)
g <- graph_from_data_frame(norm_GRN_tf_Rbpj, directed=TRUE)
edge_list <- as_edgelist(g)
g <- graph.data.frame(norm_GRN_tf_Rbpj, directed=TRUE)
weights <- sapply(E(g), function(e) {norm_GRN_tf_Rbpj$prob[which(norm_GRN_tf_Rbpj$tf == V(g)[e]$name & norm_GRN_tf_Rbpj$target == V(g)[e]$name)]})

cat("Number of edges in graph:", ecount(g), "\n")
cat("Length of weights vector:", length(weights), "\n")

cat("Node names in graph:", head(V(g)$name), "\n")
cat("Unique values in 'tf' column of 'norm_GRN_tf_Rbpj':", head(unique(norm_GRN_tf_Rbpj$tf)), "\n")
cat("Unique values in 'target' column of 'norm_GRN_tf_Rbpj':", head(unique(norm_GRN_tf_Rbpj$target)), "\n")

cat("Nodes in graph but not in 'norm_GRN_tf_Rbpj':", setdiff(V(g)$name, unique(norm_GRN_tf_Rbpj$target)), "\n")
```

# Creating input GRN for CYTOSCAPE
## (old version)
Too many edges are represented with this code because the threshold is always arbitrary
```{r}

target_names <- unique(GRN$target)

# Check that all target names are in the graph
cat("Target nodes in graph:", head(target_names[target_names %in% V(g)$name]), "\n")

# Check that all graph nodes are in the target names
cat("Graph nodes in target names:", head(V(g)$name[V(g)$name %in% target_names]), "\n")

# Extract weights from data frame
# Extract weights from data frame
# Extract weights from data frame
weights <- sapply(E(g), function(e) {
  tf_name <- V(g)[tail(e)]$name
  target_name <- V(g)[head(e)]$name
  prob <- norm_GRN_tf_Rbpj$prob[norm_GRN_tf_Rbpj$tf == tf_name & norm_GRN_tf_Rbpj$target == target_name]
  if (length(prob) == 0) {
    warning(paste("No weight found for edge", tf_name, "->", target_name))
    prob <- 0
  }
  prob
})
sum(weights)


# Check that weights were extracted correctly
cat("Sample weights:", head(weights), "\n")

```



## Expression data: DEGs against all Liver Cell Atlas
https://www.cell.com/cell/pdf/S0092-8674(21)01481-1.pdf
```{r}
# load the data
cell_types <- c("KC","HSC","LSEC","Hep","Neutro")
# define Min-Max normalization function
min_max_norm <- function(x) {(x - min(x)) / (max(x) - min(x))}

for (cell in cell_types){
  if (cell == "HSC"){
    DEGs <- read.table(paste0("../Data/ATLAS/MouseStSt/DEGsAtlas/DEGsCD45neg_mouse_",cell,".tsv"), header = TRUE, sep = "\t")
  }else{
    DEGs <- read.table(paste0("../Data/ATLAS/MouseStSt/DEGsAtlas/DEGsAll_mouse_",cell,".tsv"), header = TRUE, sep = "\t")
  }
  DEGs$score_propdeANDlfc <- rowMeans(data.frame(min_max_norm(DEGs$lfc_mean),min_max_norm(DEGs$proba_de)))
  DEGs <- DEGs[order(DEGs$score_propdeANDlf, decreasing = TRUE),]
  rownames(DEGs) <- seq(1:dim(DEGs)[1])
  colnames(DEGs)[1] <- "gene"
  DEGs <- DEGs[c('gene','score_propdeANDlfc', 'proba_de', 'lfc_mean')]
  
  slot(NICHE_qPCR,cell)[["DEGs"]] <- DEGs
}
NICHE_qPCR@KC[["DEGs"]]
```



## GRN pre-processing for visualization (new version)
First, for the core gene (e.g. Rbpj) we take the 10 out-edges. We take the top 10-5 genes for each 1st level node. For TFs we take the top 5 TFs controlling its expression and the top 5 genes they control.
```{r}
cell = "KC"
GRN <- slot(NICHE_qPCR,cell)[["GRN"]]
dim(GRN)
head(GRN)
core_node = "Rbpj" # select the core gene for the GRN visualization

allTFs <- unique(sort(GRN$tf))
alltargets <- unique(sort(GRN$target))

length(allTFs)
length(alltargets)
length(allTFs)*length(alltargets)
dim(GRN)
# add extra column indicating if the target_gene is a TF or a normal gene
GRN$is.target.a.TF <- ifelse(GRN$target %in% allTFs, "yes", "no")
# Normalize the TF prob values by TF group
library(dplyr)
GRN_TFnorm <- GRN %>%
  group_by(tf) %>%
  mutate(prob = (prob - min(prob)) / (max(prob) - min(prob))) %>%
  ungroup() 

GRN_TFnorm <- GRN_TFnorm[order(GRN_TFnorm$prob, decreasing = TRUE),]

summary(GRN_TFnorm[GRN_TFnorm$tf == core_node,])
GRN_TFnorm[GRN_TFnorm$tf == core_node,]
```

Now we calculate the DEG score combining the probability of the edge with the probability of the target gene to be differential expressed (DE) from respect to all the liver cells and the log2FC expression respect all those cells:
```{r}
# DEGs score
DEGs <- slot(NICHE_qPCR,cell)[["DEGs"]]
GRN_TFnorm$target.propdeANDlfc <- ifelse(GRN_TFnorm$target %in% DEGs$gene, DEGs$score_propdeANDlfc[match(GRN_TFnorm$target, DEGs$gene)], 0)

# create a score between probability of ANANSE and DEGs
GRN_TFnorm$score.DEGsel <- rowMeans(GRN_TFnorm[, c("prob", "target.propdeANDlfc")])
GRN_TFnorm <- GRN_TFnorm[order(GRN_TFnorm$score.DEGsel, decreasing = TRUE), ]

# we normalize the score.DEGsel per TF group
library(dplyr)
GRN_TFnorm <- GRN_TFnorm %>%
  group_by(tf) %>%
  mutate(score.DEGsel = (score.DEGsel - min(score.DEGsel)) / (max(score.DEGsel) - min(score.DEGsel))) %>%
  ungroup() 

summary(GRN_TFnorm[GRN_TFnorm$tf == core_node,])
GRN_TFnorm[GRN_TFnorm$tf == core_node,]
```

```{r}
# create metadata table with info of which gene is TF and which one not
metadata.GRN_TFnorm <- data.frame(gene = alltargets, 
                 is.TF = ifelse(alltargets %in% allTFs, "yes", "no"),
                 target.propdeANDlfc = ifelse(alltargets %in% DEGs$gene, DEGs$score_propdeANDlfc[match(alltargets, DEGs$gene)], 0))

metadata.GRN_TFnorm <- metadata.GRN_TFnorm[order(metadata.GRN_TFnorm$is.TF, metadata.GRN_TFnorm$target.propdeANDlfc),]
head(metadata.GRN_TFnorm)
tail(metadata.GRN_TFnorm)

```

## 1. TFs-->Core-->Genes NETWORK
First, we create the network with top 10 TFs controlling the core_node / core TF (e.g. Rbpj) and the top 10 genes controlled by this core TF.
```{r}
dim(GRN_TFnorm)
topA = 15
# genes whose expression is activated by the core TF
from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node,]
from_core <- from_core[order(from_core$prob, decreasing = TRUE)[1:topA],]
from_core
# TFs activating the core TF expression
to_core <- GRN_TFnorm[GRN_TFnorm$target == core_node,]
to_core <- to_core[order(to_core$prob, decreasing = TRUE)[1:topA],]
to_core
# Edge of the core TF activating himself
auto_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$target == core_node,]
auto_core
# Rest of the edges connecting the all the nodes
all_nodes <- unique(c(core_node, from_core$target, to_core$tf))
all_nodes
cytoGRN.X <- GRN_TFnorm[GRN_TFnorm$tf %in% all_nodes & GRN_TFnorm$target %in% all_nodes,]
cytoGRN.X <- cytoGRN.X[order(cytoGRN.X$prob, decreasing = TRUE),]
cytoGRN.X
message("Total number of nodes: ",length(all_nodes))
message("Total number of edges: ",dim(cytoGRN.X)[1])
```

```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN.X, paste0("../Results/Cytoscape/",cell,"/cytoGRN_X_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
```

## 2. Core-->DownGenes NETWORK
### 2.1. Without DEGs filtering
Second, we create the network with top 15 genes/TFs controlled by the core_node / core TF (e.g. Rbpj) and if they are genes we show the top 3 TF controlling this gene, and if it is TF we show the top 5 genes controlled by this TF.
```{r}
dim(GRN_TFnorm)
topA = 10
topB = 10
topC = 3
# genes whose expression is activated by the core TF
from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node,]
from_core <- from_core[order(from_core$prob, decreasing = TRUE)[1:topA],]
from_core
tfs_from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$is.target.a.TF == "yes",]
tfs_from_core <- tfs_from_core[order(tfs_from_core$prob, decreasing = TRUE)[1:topB],]
tfs_from_core
# TFs activating the core TF expression
tfs_genes <- NULL
for (TF in tfs_from_core$target){
  TF_genes <- GRN_TFnorm[GRN_TFnorm$tf == TF,]
  TF_genes <- TF_genes[order(TF_genes$prob, decreasing = TRUE)[1:topC],]
  tfs_genes <- rbind(tfs_genes, TF_genes)
}
tfs_genes
# Edge of the core TF activating himself
auto_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$target == core_node,]
auto_core
# Rest of the edges connecting the all the nodes
all_nodes <- unique(c(core_node, from_core$target, tfs_from_core$target, tfs_genes$target))
all_nodes
cytoGRN.D <- GRN_TFnorm[GRN_TFnorm$tf %in% all_nodes & GRN_TFnorm$target %in% all_nodes,]
cytoGRN.D <- cytoGRN.D[order(cytoGRN.D$prob, decreasing = TRUE),]
cytoGRN.D
message("Total number of nodes: ",length(all_nodes))
message("Total number of edges: ",dim(cytoGRN.D)[1])
```

```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN.DT, paste0("../Results/Cytoscape/",cell,"/cytoGRN_DT_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
write.table(cytoGRN.DTjust, paste0("../Results/Cytoscape/",cell,"/cytoGRN_DTjust_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
core_node
```

## 2.2. With DEGs filtering
Second, we create the network with top 15 genes/TFs controlled by the core_node / core TF (e.g. Rbpj) and if they are genes we show the top 3 TF controlling this gene, and if it is TF we show the top 5 genes controlled by this TF.
```{r}
dim(GRN_TFnorm)
topA = 100
topB = 50
topC = 60
# genes whose expression is activated by the core TF
#
from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node,]
from_core <- from_core[order(from_core$score.DEGsel, decreasing = TRUE)[1:topA],]
from_core
tfs_from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$is.target.a.TF == "yes",]
tfs_from_core <- tfs_from_core[order(tfs_from_core$score.DEGsel, decreasing = TRUE)[1:topB],]
tfs_from_core
auto_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$target == core_node,]

# TFs activating the core TF expression
tfs_genes <- NULL
for (TF in c(tfs_from_core$target)){
  TF_genes <- GRN_TFnorm[GRN_TFnorm$tf == TF,]
  TF_genes <- TF_genes[order(TF_genes$score.DEGsel, decreasing = TRUE)[1:topC],]
  tfs_genes <- rbind(tfs_genes, TF_genes)
}
tfs_genes

# Rest of the edges connecting the all the nodes
all_nodes <- unique(c(core_node, from_core$target, tfs_from_core$target, tfs_genes$target, 'Plvap'))
all_nodes
cytoGRN.Ddeg <- GRN_TFnorm[GRN_TFnorm$tf %in% all_nodes & GRN_TFnorm$target %in% all_nodes,]
cytoGRN.Ddeg <- cytoGRN.Ddeg[order(cytoGRN.Ddeg$score.DEGsel, decreasing = TRUE),]
cytoGRN.Ddeg
message("Total number of nodes: ",length(all_nodes))
message("Total number of edges: ",dim(cytoGRN.Ddeg)[1])
# without upgoing edges, just downstream edges
cytoGRN.Ddegjust <- rbind(from_core, tfs_from_core, auto_core, tfs_genes)
cytoGRN.Ddegjust <- cytoGRN.Ddegjust[order(cytoGRN.Ddegjust$score.DEGsel, decreasing = TRUE),]
cytoGRN.Ddegjust
```

```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN.Ddeg, paste0("../Results/Cytoscape/",cell,"/cytoGRN_Ddeg_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
write.table(cytoGRN.Ddegjust, paste0("../Results/Cytoscape/",cell,"/cytoGRN_Ddegjust_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
core_node
```


## 3. Downstream GRN with TFS with DEGs filtering
```{r}
dim(GRN_TFnorm)
topA = 10
topB = 15
topC = 20
# TF whose expression is activated by the core TF
tfs_from_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node & GRN_TFnorm$is.target.a.TF == "yes",]
x <- tfs_from_core$score.DEGsel
tfs_from_core$score.DEGsel <- (x - min(x))/(max(x) - min(x))
auto_core <- tfs_from_core[tfs_from_core$tf == core_node & tfs_from_core$target == core_node,] # Edge of the core TF activating himself
tfs_from_core <- tfs_from_core[order(tfs_from_core$score.DEGsel, decreasing = TRUE)[1:topA],]
tfs_from_core

# TFs activating the core TF expression
tfs_genes <- NULL
for (TF in tfs_from_core$target){
  TF_genes <- GRN_TFnorm[GRN_TFnorm$tf == TF,]
  TF_genes <- TF_genes[order(TF_genes$score.DEGsel, decreasing = TRUE)[1:topC],]
  tfs_genes <- rbind(tfs_genes, TF_genes)
}
tfs_genes

# Rest of the edges connecting the all the nodes
all_nodes <- unique(c(core_node, tfs_from_core$target, tfs_genes$target))
all_nodes
cytoGRN.DTdeg <- GRN_TFnorm[GRN_TFnorm$tf %in% all_nodes & GRN_TFnorm$target %in% all_nodes,]
cytoGRN.DTdeg <- cytoGRN.DTdeg[order(cytoGRN.DTdeg$score.DEGsel, decreasing = TRUE),]
#cytoGRN.DTdeg <- cytoGRN.DTdeg[cytoGRN.DTdeg$score.DEGsel > 0.96,]
cytoGRN.DTdeg
message("Total number of nodes: ",length(all_nodes))
message("Total number of edges: ",dim(cytoGRN.DT)[1])
# without upgoing edges, just downstream edges
cytoGRN.DTdegjust <- rbind(tfs_from_core,auto_core,tfs_genes)
cytoGRN.DTdegjust <- cytoGRN.DTdegjust[order(cytoGRN.DTdegjust$score.DEGsel, decreasing = TRUE),]
cytoGRN.DTdegjust
# cytoGRN.DTdegjust <- cytoGRN.DTdegjust[cytoGRN.DTdegjust$score.DEGsel > 0.96,]
```

```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN.DTdeg, paste0("../Results/Cytoscape/",cell,"/cytoGRN_DTdeg_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
write.table(cytoGRN.DTdegjust, paste0("../Results/Cytoscape/",cell,"/cytoGRN_DTdegjust_",cell,"_",core_node,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
core_node
```

# 4. GRN with 5% of edges with highest probability values
```{r}
round(dim(GRN)[1]*0.05)
cytoGRN.05 <- GRN[order(-GRN$prob),][1:round(dim(GRN)[1]*0.05),]
cytoGRN.05
```


```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN.05, paste0("../Results/Cytoscape/",cell,"/cytoGRN_05_",cell,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
core_node
```


# 5. M.Guilliams network: top 5 DEGs TFs, top 10 DEGs targets, (and top 5 DEGs controlled by the other TFs)
```{r}
dim(DEGs)
DEGs
GRN_TFnorm_core <- GRN_TFnorm[GRN_TFnorm$tf == core_node,]
GRN_MG_core <- GRN_TFnorm_core[GRN_TFnorm_core$target %in% DEGs$gene,]
GRN_MG_core # all the target genes in this network are in the DEGs list
TFs_core <- GRN_MG_core[GRN_MG_core$is.target.a.TF == "yes",]
TFs_core <- TFs_core[order(-TFs_core$score.DEGsel),]
TFs_core
targets_core <- GRN_MG_core[GRN_MG_core$is.target.a.TF == "no",]
targets_core <- targets_core[order(-targets_core$score.DEGsel),]
targets_core
cytoGRN_MG <- rbind(TFs_core[1:7,],targets_core[1:10,])
cytoGRN_MG <- cytoGRN_MG[order(-cytoGRN_MG$score.DEGsel),]
cytoGRN_MG
```


```{r}
# Write the GRN dataframe to a tsv file to represent in Cytoscape
write.table(cytoGRN_MG, paste0("../Results/Cytoscape/",cell,"/cytoGRN_MG_",core_node,"_",cell,".tsv"), sep = "\t", quote = F, col.names = NA, row.names = TRUE)
core_node
```
